### 2018-01-30 experiments/3_processed_data/mturk_data_processing.Rmd
added the methods explanations

---
title: "mturk_data_processing"
author: "Masoud Jasbi"
date: "3/7/2017"
output: html_document
---

```{r cars}
library(rjson)
library(tidyverse)
```

```{r}
# read the json file names to be loaded in the json directory
files <- dir("../2_raw_data/", pattern =".json") 
# make an empty data frame to append the json data participant by participant
adult_data <- data.frame()

#loop through json files
for (file in files) {
  #read the json file into jd
  jd <- fromJSON(paste(readLines(paste("../2_raw_data/", file, sep="")), collapse=""))
  #create a data frame with the column names on the left
  id <- data_frame(participant = jd$WorkerId,
                   order=seq(1:24),
                  gender = jd$answers$data$gen,
                  age = jd$answers$data$age,
                  language = jd$answers$data$language,
#                  version = jd$answers$data$version,
                  logical_training = jd$answers$data$education,
                  trials = jd$answers$data$trial,
                  response = jd$answers$data$response,
#                  duration = jd$answers$data$elapsed_ms,
#                  AgeGroup = "adult",
                  aim = jd$answers$data$expt_aim,
                  comments = jd$answers$data$expt_gen
                  )
  # append the dataframe for that json file to the big data frame 
  adult_data <- bind_rows(adult_data, id)
}
write.csv(adult_data, "../3_processed_data/adult_data.csv")
```

```{r}
processed_data <- read.csv("../3_processed_data/adult_data_withConditions.csv")

processed_data2 <-
  processed_data %>%
  separate(trials, c("card_type", "guess_type", "card", "guess"), sep = "_", extra = "drop")

write.csv(processed_data2, "../3_processed_data/processed_data2.csv")

processed_data2 %>%
  group_by(type, card_type, guess_type) %>%
  summarize(count=n())
```