### 2018-02-23 experiments/pilot/3_processed_data/processed_data_processing.Rmd
before pulling the final website changes

---
title: "mturk_data_processing"
author: "Masoud Jasbi"
date: "3/7/2017"
output: html_document
---

```{r cars}
library(tidyverse)
```

```{r}
data <- read.csv("adult_data_withConditions.csv")

data[data$type=="binary" & data$response==0,]$response <- "wrong"
data[data$type=="binary" & data$response==1,]$response <- "right"

data[data$type=="ternary" & data$response==0,]$response <- "wrong"
data[data$type=="ternary" & data$response==1,]$response <- "right"
data[data$type=="ternary" & data$response==2,]$response <- "neiter"

data[data$type=="quatenary" & data$response==0,]$response <- "wrong"
data[data$type=="quatenary" & data$response==1,]$response <- "kinda wrong"
data[data$type=="quatenary" & data$response==2,]$response <- "kinda right"
data[data$type=="quatenary" & data$response==3,]$response <- "right"

final_data <- 
  data %>% 
  filter(type!="slider") %>% 
  separate(trials, into = c("card_type", "guess_type", "card", "guess"), sep = "_") %>%
  unite(trial_type, card_type, guess_type, remove = FALSE) %>%
  rename(response_type = "type")

write.csv(final_data, "guessingGame_data.csv", row.names = FALSE)
```