### 2018-02-24 experiments/pilot/1_methods/methods.rmd
pilot 2 data

---
title: "Card Game Methods"
author: "Masoud Jasbi"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---

### Card Types:

There are three animals in the card-game: cat, dog, and elephant. Each card may have one or two of them. I represent the card types in the following way:

| Representation | Description | Cards |
|---------|-----------|---------------------------|
|**X**| Only one animal on the card |  <img src="stimuli/oneanimal_card.jpg" width="300" height="300" /> |
|**XY**| Two animals on the card | <img src="stimuli/twoanimal_card.jpg" width="300" height="300" /> |

## Guess Types

| Representation | Description | 
|---|------------------------------------------------------------|
| **Z** | The puppet mentions an animal not on the card | 
| **X** | The puppet mentions one of the animals (perhaps the only animal) on the card |
| **XorY** | The puppet mentions both animals on the card and uses the connective "or" |
| **XandY** | The puppet mentions both animals on the card and uses the connective "and"|

## Trial Types

The combination of card types and guess types creats 12 possible trial types and 36 possible trials. We choose 8 trial types out of 12 to balance for the type of card (X vs. XY), the type of guess (with vs. without connective), and types of responses (wrong, right). We represent each trial type using the pattern "CARD-GUESS" with X, Y and Z abstracting over the specific animals:

| Representation | Example Card | Example Guess |
|---|---|------------------------|
|  **XY-XorY** | <img src="stimuli/catdog_card.jpg" width="70" height="70" /> | There is a cat **or** a dog |
| **XY-XandY** | <img src="stimuli/catdog_card.jpg" width="70" height="70" />| There is a cat **and** a dog |
| **XY-X** | <img src="stimuli/catdog_card.jpg" width="70" height="70" />| There is a cat! |
| **XY-Z** | <img src="stimuli/catdog_card.jpg" width="70" height="70" />| There is an elephantl! |
| **X-XorY** | <img src="stimuli/cat_card.jpg" width="70" height="70" />| There is a cat or a dog! |
| **X-XandY** | <img src="stimuli/cat_card.jpg" width="70" height="70" />| There is a cat and a dog! |
|  **X-X** | <img src="stimuli/cat_card.jpg" width="70" height="70" />| There is a cat! |
| **X-Z** | <img src="stimuli/cat_card.jpg" width="70" height="70" />| There is a dog! |

Below, I create trial type variables that hold lists of the trials in each trial type:
```{r}
### Two animals, two correct lables with or:
XY.XorY <- list(c("CAT-DOG", "cat or dog"), c("CAT-ELE", "cat or elephant"), c("DOG-ELE", "dog or elephant"))

### Two animals, two correct lables with or: 
XY.XandY <- list(c("CAT-DOG", "cat and dog"), c("CAT-ELE", "cat and elephant"), c("DOG-ELE", "dog and elephant"))

### Two animals, one correct lable:
XY.X <- list(c("CAT-DOG", "cat"), c("CAT-DOG", "dog"), c("CAT-ELE", "cat"), c("CAT-ELE", "elephant"), c("DOG-ELE", "dog"), c("DOG-ELE", "elephant"))

### Two animals, one incorrect lable:
XY.Z <- list(c("CAT-DOG", "elephant"), c("CAT-ELE", "dog"), c("DOG-ELE", "cat"))

### One animal, two lables with or, one label correct
X.XorY <- list(c("CAT", "cat or dog"), c("DOG", "cat or dog"), c("CAT", "cat or elephant"), c("ELE", "cat or elephant"), c("DOG", "dog or elephant"), c("ELE", "dog or elephant"))

### One animal, two lables with and, one label correct
X.XandY <- list(c("CAT", "cat and dog"), c("DOG", "cat and dog"), c("CAT", "cat and elephant"), c("ELE", "cat and elephant"), c("DOG", "dog and elephant"), c("ELE", "dog and elephant"))

### one animal, one correct label:
X.X <- list(c("CAT", "cat"), c("DOG", "dog"), c("ELE", "elephant"))

### one animal, one incorrect label:
X.Z <- list(c("CAT", "dog"), c("CAT", "elephant"), c("DOG", "cat"), c("DOG", "elephant"), c("ELE", "cat"), c("ELE", "dog"))
```

### Randomization:

With 8 trial types, it is reasonable to assume that children can sit through two trials per trial type: 16 trials total. Therefore we need to randomly sample 2 trials from the trial types for each participant. Some trial types have 6 trials and some 3. For those with 6 trials I only randomize them. For those with 3, I randomize them twice and bind them to create a list of six trials. This way, each trial type has 6 randomized trials.

```{r}
XY.XorY.s <- c(sample(XY.XorY), sample(XY.XorY))
XY.XandY.s <- c(sample(XY.XandY), sample(XY.XandY))
XY.X.s <- sample(XY.X)
XY.Z.s <- c(sample(XY.Z), sample(XY.Z))
X.XorY.s <- sample(X.XorY)
X.XandY.s <- sample(X.XandY)
X.X.s <- c(sample(X.X), sample(X.X)) 
X.Z.s <- sample(X.Z)
```

Now for each trial type, I pick the first two trials for the first participant, the second two for the second, and the third two for the third.

Now there are two main bocks. Each block has

```{r}
# Taking the first and the second trial of each trial type for participant 1:
part1 <- c(
  sample(c(X.Z.s[1], X.X.s[1], XY.X.s[1])), 
  sample(c(XY.XorY.s[1], XY.XandY.s[1], XY.Z.s[1], X.XorY.s[1], X.XandY.s[1])),
  sample(c(XY.XorY.s[2], XY.XandY.s[2], XY.X.s[2], XY.Z.s[2], X.XorY.s[2], X.XandY.s[2], X.X.s[2], X.Z.s[2]))
          )
# Taking the third and the fourth trial of each trial type for participant 2:
part2 <- c(
  sample(c(X.Z.s[3], X.X.s[3], XY.X.s[3])), 
  sample(c(XY.XorY.s[3], XY.XandY.s[3], XY.Z.s[3], X.XorY.s[3], X.XandY.s[3])),
  sample(c(XY.X.s[4], XY.XorY.s[4], XY.XandY.s[4], XY.Z.s[4], X.XorY.s[4], X.XandY.s[4], X.X.s[4], X.Z.s[4]))
        )

# Taking the fifth and the sixth trial of each trial type for participant 1:
part3 <- c(
  sample(c(X.Z.s[5], X.X.s[5], XY.X.s[5])), 
  sample(c(XY.XorY.s[5], XY.XandY.s[5], XY.Z.s[5], X.XorY.s[5], X.XandY.s[5])),
  sample(c(XY.XorY.s[6], XY.XandY.s[6], XY.X.s[6], XY.Z.s[6], X.XorY.s[6], X.XandY.s[6], X.Z.s[6], X.X.s[6]))
  )
```

The code below saves randomized trials for 3 participants in the expeirment in three separate CSV files.

```{r eval=FALSE}
######### Creating the CSV files:

## File 1:
file1 <- c()
for (i in 1:16) {
   file1 <- rbind(file1, part1[[i]])
}
colnames(file1) <- c("Card", "Guess")
write.csv(file1, file="file1.csv", row.names=FALSE)

## File 2:
file2 <- c()
for (i in 1:16) {
   file2 <- rbind(file2, part2[[i]])
}

colnames(file2) <- c("Card", "Guess")
write.csv(file2, file="file2.csv", row.names=FALSE)

## File 3:
file3 <- c()
for (i in 1:16) {
   file3 <- rbind(file3, part3[[i]])
}

colnames(file3) <- c("Card", "Guess")
write.csv(file3, file="file3.csv", row.names=FALSE)
```