### 2018-04-12 experiments/main/3_processed_data/mturk_processing.Rmd
data collection and plotting

---
title: "mturk_data_processing"
author: "Masoud Jasbi"
date: "3/7/2017"
output: html_document
---

```{r cars}
library(rjson)
library(tidyverse)
```

```{r}
# read the json file names to be loaded in the json directory
files <- dir("../2_raw_data/", pattern =".json") 
# make an empty data frame to append the json data participant by participant
adult_data <- data.frame()

#loop through json files
for (file in files) {
  #read the json file into jd
  jd <- fromJSON(paste(readLines(paste("../2_raw_data/", file, sep="")), collapse=""))
  #create a data frame with the column names on the left
  id <- data_frame(participant = jd$WorkerId,
                   order=seq(1:24),
                  gender = jd$answers$data$gen,
                  age = jd$answers$data$age,
                  language = jd$answers$data$language,
                  response_type = jd$answers$data$response_type,
                  trial_type = jd$answers$data$trial_type,
                  guess_type = jd$answers$data$guess_type,
                  guess = jd$answers$data$guess,
                  card_type = jd$answers$data$card_type,
                  card = jd$answers$data$card,
                  response = jd$answers$data$response,
                  logical_training = jd$answers$data$logical_training,
                  duration = jd$answers$data$elapsed_ms,
                  aim = jd$answers$data$aim,
                  comments = jd$answers$data$comments
                  )
  # append the dataframe for that json file to the big data frame 
  adult_data <- bind_rows(adult_data, id)
}
write_csv(adult_data, "../3_processed_data/data_main.csv")
```