### 2018-12-19 manuscript.tex
second revisions

\documentclass[man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

  \usepackage{ifthen} % Only add declarations when endfloat package is loaded
  \ifthenelse{\equal{\string man}{\string man}}{%
   \DeclareDelayedFloatFlavor{ThreePartTable}{table} % Make endfloat play with longtable
   % \DeclareDelayedFloatFlavor{ltable}{table} % Make endfloat play with lscape
   \DeclareDelayedFloatFlavor{lltable}{table} % Make endfloat play with lscape & longtable
  }{}%

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

  \usepackage{graphicx}
  \makeatletter
  \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
  \def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
  \makeatother
  % Scale images if necessary, so that they will not overflow the page
  % margins by default, and it is still possible to overwrite the defaults
  % using explicit options in \includegraphics[width, height, ...]{}
  \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Linking hypothesis and number of response options modulate inferred scalar implicature rate},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{color}
\definecolor{Blue}{RGB}{0,0,250}
\newcommand{\change}[1]{\textcolor{Blue}{#1}}

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Linking hypothesis and number of response options modulate inferred
scalar implicature rate}

  \shorttitle{Linking hypotheses and implicature rate}

  \author{Masoud Jasbi\textsuperscript{1}, Brandon Waldon\textsuperscript{2}, \& Judith Degen\textsuperscript{2}}

  % \def\affdep{{"", "", ""}}%
  % \def\affcity{{"", "", ""}}%

  \affiliation{
    \vspace{0.5cm}
       \textsuperscript{1} Harvard University, Department of Linguistics\\ 
	\textsuperscript{2} Stanford University, Department of Linguistics }
  \authornote{
    The authors declare that they have no affiliations with, or involvement
    in any organization or entity with any financial interest, or
    non-financial interest in the subject matter or materials discussed in
    this manuscript.

    Correspondence concerning this article should be addressed to Masoud
    Jasbi, Department of Linguistics, Boylston Hall, Cambridge, MA,
    94305-2150. E-mail:
    \href{mailto:masoud_jasbi@fas.harvard.edu}{\nolinkurl{masoud_jasbi@fas.harvard.edu}}
  }

  \abstract{The past 15 years have seen increasing experimental investigations of
core pragmatic questions in the ever more active and lively field of
experimental pragmatics. Within experimental pragmatics, many of the
core questions have relied on the operationalization of the theoretical
notion of `implicature rate'. Implicature rate based results have
informed the work on acquisition, online processing, and scalar
diversity, inter alia.  Implicature
rate has typically been quantified as the proportion of `pragmatic'
judgments in two-alternative forced choice truth value judgment tasks.
Despite its theoretical importance, this linking hypothesis from
implicature rate to behavioral responses has never been extensively
tested. Here we show that two factors dramatically affect the
`implicature rate' inferred from truth value judgment tasks: a) the
number of responses provided to participants; and b) the linking
hypothesis about what constitutes a `pragmatic' judgment. We argue that
it is time for the field of experimental pragmatics to engage more
seriously with its foundational assumptions about how theoretical
notions map onto behaviorally measurable quantities, and present a
sketch of an alternative linking hypothesis that derives behavior in
truth value judgment tasks from probabilistic utterance expectations.}
  \keywords{scalar implicature; methodology; linking hypothesis; experimental
pragmatics; truth value judgment task \\

    \indent Word count: 9037
  }

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

\section{Introduction}\label{introduction}

The past 15 years have seen the rise and development of a bustling and
exciting new field at the intersection of linguistics, psychology, and
philosophy: \emph{experimental pragmatics} (Barner, Brooks, \& Bale,
2011; Bonnefon, Feeney, \& Villejoubert, 2009; Bott \& Chemla, 2016;
Bott \& Noveck, 2004; Breheny, Ferguson, \& Katsos, 2013; Breheny,
Katsos, \& Williams, 2006; Chierchia et al., 2001; De Neys \& Schaeken,
2007; Degen \& Tanenhaus, 2015, 2016; Geurts \& Pouscoulous, 2009;
Grodner, Klein, Carbary, \& Tanenhaus, 2010; Huang \& Snedeker, 2009;
Katsos \& Bishop, 2011; I. A. Noveck \& Reboul, 2008; Noveck \& Posada,
2003; Papafragou \& Tantalou, 2004; van Tiel, van Miltenburg, Zevakhina, \&
Geurts, 2016; Tomlinson, Bailey, \& Bott, 2013). Experimental pragmatics
is devoted to experimentally testing theories of how language is used in
context. How do listeners draw inferences about the -- often
underspecified -- linguistic signal they receive from speakers? How do
speakers choose between the many utterance alternatives they have at
their disposal?

The most prominently studied phenomenon in experimental pragmatics is
undoubtedly \emph{scalar implicature}. Scalar implicatures arise as a
result of a speaker producing the weaker of two ordered scalemates
(Geurts, 2010; Grice, 1975; Hirschberg, 1985; Horn, 1972). Examples are
provided in (1-2).

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\tightlist
\item
  Some of her pets are cats.
\end{enumerate}

\emph{Implicature:} Some, but not all, of her pets are cats.

\emph{Scale:} $\langle$all, some$\rangle$

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  She owns a cat or a dog.
\end{enumerate}

\emph{Implicature:} She owns a cat or a dog, but not both.

\emph{Scale:} $\langle$and, or$\rangle$

A listener, upon observing the utterances in (1-2) typically infers
that the speaker intended to convey the meanings listed as \emph{Implicature}s,
respectively. Since Grice (1975), the agreed-upon abstract
rationalization the listener could give for their inference goes
something like this: the speaker could have made a more informative
statement by producing the stronger alternative (e.g., \emph{All of her
pets are cats} in (1)). If the stronger alternative is true, they
should have produced it to comply with the Cooperative Principle. They
chose not to. Assuming the speaker knows whether the stronger
alternative is true, it must not be true. The derivation
procedure for ad hoc exhaustivity inferences such as in (3) is assumed to
be calculable in the same way as for scalar implicatures, though the
scale is assumed to be contextually driven.

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  She owns a cat.
\end{enumerate}

\emph{Implicature:} She owns only a cat.

\emph{Scale:} $\langle$cat and dog, cat$\rangle$\\
~\\

Because the basic reconstruction of the inference is much more easily
characterized for scalar implicatures than for other implicatures,
scalar implicatures have served as a test bed for many questions in
experimental pragmatics, including, but not limited to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Are scalar inferences default inferences, in the sense that they arise
  unless blocked by (marked) contexts (Degen, 2015; Horn, 1984;
  Levinson, 2000)?
\item
  Are scalar inferences default inferences, in the sense that they are
  computed automatically in online processing and only cancelled in a second effortful step if required by context (Bott \&
  Noveck, 2004; Breheny et al., 2006; Degen \& Tanenhaus, 2016; Grodner
  et al., 2010; Huang \& Snedeker, 2009; Politzer-Ahles \& Fiorentino,
  2013; Tomlinson et al., 2013)?
\item
  What are the (linguistic and extra-linguistic) factors that affect
  whether a scalar implicature is derived (Bergen \& Grodner, 2012;
  Bonnefon et al., 2009; Breheny et al., 2013, 2006; Chemla \& Spector,
  2011; De Neys \& Schaeken, 2007; Degen, 2015; Degen \& Goodman, 2014;
  Degen \& Tanenhaus, 2015, 2016; Marneffe \& Tonhauser, 2016; Potts,
  Lassiter, Levy, \& Frank, 2015; Zondervan, 2010)?
\item
  How much diversity is there across implicature types, and within
  scalar implicatures across scale types, in whether or not an
  implicature is computed (Doran, Ward, Larson, McNabb, \& Baker, 2012;
  van Tiel et al., 2016)?
\item
  At what age do children acquire the ability to compute implicatures
  (Barner et al., 2011; Horowitz, Schneider, \& Frank, 2017; Katsos \&
  Bishop, 2011; Musolino, 2004; Noveck, 2001; Papafragou \& Tantalou,
  2004; Stiller, Goodman, \& Frank, 2015)?
\end{enumerate}

In addressing all of these questions, it has been important to obtain
estimates of \emph{implicature rates}. For 1., implicature rates from
experimental tasks can be taken to inform whether scalar implicatures
should be considered default inferences. For 2., processing measures on
responses that indicate implicatures can be compared to processing
measures on responses that indicate literal interpretations. For 3.,
contextual effects can be examined by comparing implicature rates across
contexts. For 4., implicature rates can be compared across scales (or
across implicature types). For 5., implicature rates can be compared
across age groups.

A standard measure that has stood as a proxy for implicature rate across
many studies is the proportion of \enquote{pragmatic} judgments in truth
value judgment paradigms (Bott \& Noveck, 2004; Chemla \& Spector, 2011;
De Neys \& Schaeken, 2007; Degen \& Goodman, 2014; Degen \& Tanenhaus,
2015; Geurts \& Pouscoulous, 2009; Noveck, 2001; Noveck \& Posada,
2003). In these kinds of tasks, participants are provided a set of
facts, either presented visually or via their own knowledge of the
world. They are then asked to judge whether a sentence intended to
describe those facts is true or false (or alternatively, whether it is
right or wrong, or they are asked whether they agree or disagree with
the sentence). The crucial condition for assessing implicature rates in
these kinds of studies typically consists of a case where the facts are
such that the stronger alternative is true and the target utterance is
thus also true but underinformative. For instance, Bott and Noveck
(2004) asked participants to judge sentences like \enquote{Some
elephants are mammals}, when world knowledge dictates that all elephants
are mammals. Similarly, Degen and Tanenhaus (2015) asked participants to
judge sentences like \enquote{You got some of the gumballs} in
situations where the visual evidence indicated that the participant
received all the gumballs from a gumball machine. In these kinds of
scenarios, the story goes, if a participant responds \enquote{FALSE},
that indicates that they computed a scalar implicature, e.g. to the effect
of \enquote{Not all elephants are mammals} or \enquote{You didn't get
all of the gumballs}, which is (globally or contextually) false. If
instead a participant responds \enquote{TRUE}, that is taken to indicate
that they interpreted the utterance literally as ``Some, and possibly
all, elephants are mammals'' or \enquote{You got some, and possibly all,
of the gumballs}.

Using the proportion of \enquote{FALSE} responses on true but underinformative trials as a proxy for implicature rate is  common in experimental pragmatics. For example, in one of the first studies to investigate scalar implicatures experimentally, Noveck (2001) tested adults' and children's interpretations of the scalar items \emph{might} and \emph{some}. The dependent measure in Noveck (2001) was the rate of ``logically correct responses'', i.e. responding ``yes'' to statements such as \emph{Some giraffes have long necks} or \emph{There might be a parrot [in the box]} when there had to be a parrot in the box. He found that children responded ``yes'' more frequently than adults, and concluded that children interpret scalar items \emph{some} and \emph{might} more logically (i.e. literally). Similarly in another landmark study, Papafragou and Musolino (2003) tested children and adults interpretation of the following set of scalar items: <\emph{two}, \emph{three}>, <\emph{some}, \emph{all}>, and <\emph{finish}, \emph{start}>. The dependent measure in this study was the proportion of ``No'' responses to a puppet's underinformative] statement. The study concluded that ``while adults overwhelmingly rejected infelicitous descriptions, children almost never did so''. Furthermore, the study compared  implicature rates across scales and concluded that ``children also differed from from adults in that their rejection rate on the numerical scale was reliably higher than on the two other scales.'' In their final experiment, Papafragou and Musolino (2003) modified their task to invite scalar inferences more easily. They reported that this manipulation resulted in a significantly higher rejection rates. Based on these results, they concluded that children's ability to compute implicatures is affected by the type of scalar item as well as children's awareness of the task's goals. Since these early pioneering studies, the rate of ``FALSE'' (or ``No'', ``Wrong'', ``Disagree'') responses on underinformative trials in truth-value judgment tasks has become a commonplace dependent measure (Doran et al (2012),  Geurts and Pouscoulous (2009), Potts, Lassiter, Levy, and Frank (2016), inter alia.)

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of ``FALSE' responses in truth value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011); Sikos et al (this issue)). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute a small, a big, or a huge strawberry to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average smaller strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

But this raises an important question: in truth value judgment tasks,
how does the researcher know whether an interpretation is literal or the
result of an implicature computation? The binary choice task typically
used is appealing in part because it allows for a direct mapping from
response options---``TRUE'' and ``FALSE'---to interpretations---literal and
pragmatic. That the seeming simplicity of this mapping is illusory
becomes apparent once a third response option is introduced, as in the
Katsos and Bishop (2011) case. How is the researcher to interpret the
intermediate option? Katsos and Bishop (2011) grouped the intermediate
option with the negative endpoint of the scale for the purpose of
categorizing judgments as literal vs.~pragmatic, i.e., they interpreted
the intermediate option as pragmatic. But it seems just as plausible
that they could have grouped it with the positive endpoint of the scale
and taken the hard line that only truly ``FALSE' responses constitute
evidence of a full-fledged implicature. The point here is that there has
been remarkably little consideration of \emph{linking hypotheses}
between behavioral measures and theoretical constructs in experimental
pragmatics, a problem in many subfields of psycholinguistics (Tanenhaus,
2004). We argue that it is time to engage more seriously with these
issues.

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
assumes two things: first, that whatever participants are doing in a
truth value judgment task, the behavioral measure can be interpreted as
providing a measure of interpretation; and second, that listeners either
do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues.
Following Degen and Goodman (2014), we will offer some remarks on why
truth value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking hypotheses.
We then discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking hypotheses in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a
depiction of either one or two animals (see Figure
\ref{fig:linkvisualization} for an example). As in a standard truth
value judgment task, the listener then observes an underinformative
utterance about this card (e.g., \enquote{There is a cat or a dog on the
card}) and is asked to provide a judgment on a scale with 2, 3, 4, or 5
response options, with endpoints \enquote{wrong} and
\enquote{right.}\footnote{An open question concerns the extent to which
  the labeling of points on the scale affects judgments (e.g.,
  \enquote{wrong}--\enquote{right} vs. \enquote{false}--\enquote{true}
  vs. \enquote{disagree}--\enquote{agree}). Studies vary in the labeling of scale points.} In the binary case, this reproduces the standard
truth value judgment task. Figure \ref{fig:linkvisualization}
exemplifies (some of) the researcher's options for grouping responses.
Under what we will call the \enquote{Strong link} assumption, only the
negative endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Under the \enquote{Weak link}
assumption, in contrast, any response that does not correspond to the
positive endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Intermediate grouping schemes are also
possible, but these are the ones we will consider here. Note that for
the binary case, the Weak and Strong link return the same categorization
scheme, but for any number of response options greater than 2, the Weak
and Strong link can in principle lead to differences in inferences about
implicature rate.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/linkvisualization-1.pdf}
\caption{\label{fig:linkvisualization}Strong and Weak link from response
options to researcher inference about scalar implicature rate,
exemplified for the disjunctive utterance when the conjunction is true.}
\end{figure}

Let's examine an example. Assume three response options (wrong, neither,
right). Assume further that each of the three responses was selected by
a third of participants, i.e., the distributions of responses is 1/3,
1/3, and 1/3. Under the Strong link, we infer that this task yielded an
implicature rate of 2/3. Under the Weak link, we infer that this task
yielded an implicature rate of 1/3. This is quite a drastic difference
if we are, for instance, interested in whether scalar implicatures are
inference defaults and we would like to interpret an implicature rate of
above an arbitrary threshold (e.g., 50\%) as evidence for such a claim.
Under the Strong link, we would conclude that scalar implicatures are
not defaults. Under the Weak link, we would conclude that they are. In
the experiment reported in the following section, we presented
participants with exactly this setup. We manipulated the number of
response options between participants and analyzed the results under
different linking hypothesis.\footnote{\change{Researchers may vary with respect to which linking hypothesis (Weak vs.~Strong, or others) they consider most plausible. Supporters of the Weak link may argue that there are three theoretically motivated categories of judgments: false, true but infelicitous, true and felicitous. Under such an account, it is plausible that false and true+felicitous responses occupy the ends of the false/true scale while true but infelicitous responses occupy the mid portion. On critical trials, participants judge underinformative statements that are true but infelicitous and therefore the mid portion of the scale can provide evidence for implicature computation. However, supporters of the Strong link may argue that if a participant computes an implicature, their response in the task should show the commitment to that interpretation by judging the underinformative utterance as false. Any other response shows that they have not truly computed an implicature. So far, these discussions have remained largely informal. In this paper we stay neutral with respect to the plausibility of each link and only aim to demonstrate the consequences of assuming them.}}

\section{Experiment}\label{experiment}

Participants played an online card game in which they were asked to
judge descriptions of the contents of cards. Different groups of
participants were presented with different numbers of response options.
On critical trials, participants were presented with descriptions for
the cards that typically result in exhaustivity implicatures
(\enquote{There is a cat on the card} when there was a cat and a dog) or
scalar implicatures (\enquote{There is a cat or a dog on the card} when
there was a cat and a dog). We categorized their responses on such
trials according to the Weak and the Strong link introduced above, and
tested whether the number of response options and the linking
hypothesis led to different conclusions about the rate of computed
implicatures in the experimental task.

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

200 participants were recruited via Amazon Mechanical Turk. They
optionally provided demographic information at the end of the study.
Participants' mean age was 35. We also asked participants if they had
any prior training in logic. 40 participants reported that they did,
while 160 had no prior training in logic. All participants' data was
included in the final analysis.\footnote{This study was carried out in accordance with the recommendations of the Common Rule, Federal Office for Human Research Protections. The protocol was approved by the Stanford University IRB 2 (non-medical research). All subjects gave Informed consent, documentation was waived by the IRB.}

\subsubsection{Materials and procedure}\label{materials-and-procedure}

The study was administered online through Amazon Mechanical
Turk.\footnote{The experiment can be viewed
  \href{https://cdn.rawgit.com/thegricean/si-paradigms/94a590f0/experiments/main/1_methods/online_experiment/connective_game.html}{here}.}
Participants were first introduced to the set of cards we used in the
study (Figure \ref{fig:stimuli}). Each card depicted one or two animals,
where an animal could be either a cat, a dog, or an elephant. Then
participants were introduced to a blindfolded fictional character called
Bob. Bob was blindfolded to avoid violations of ignorance expectations
associated with the use of disjunction (Chierchia et al., 2001;
Sauerland, 2004). Participants were told that Bob would guess the
contents of the cards and their task was to indicate whether Bob's guess
was wrong or right. On each trial, participants saw a card and a
sentence representing Bob's guess. For example, they saw a card with a
cat and read the sentence \enquote{There is a cat on the card.} They
then provided an assessment of Bob's guess. The study ended after 24
trials.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/stimuli-1.pdf}
\caption{\label{fig:stimuli}Cards used in the connective guessing game.}
\end{figure}

Two factors were manipulated within participants: card type and guess
type. There were two types of cards, cards with only one animal on them
and cards with two animals. There were three types of guesses: simple
(e.g. \emph{There is a cat}), conjunctive (e.g. \emph{There is a cat and
a dog}), and disjunctive (e.g. \emph{There is a cat or a dog}). Crossing
card type and guess type yielded trials of varying theoretical interest
(see Figure \ref{fig:trials}): critical underinformative trials that
were likely to elicit pragmatic inferences (either scalar or exhaustive)
and control trials that were either unambiguously true or false. Each
trial type occurred three times with randomly sampled animals and
utterances that satisfied the constraint of the trial type. Trial order
was randomized.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/trials-1.pdf}
\caption{\label{fig:trials}Trial types (critical and control). Headers
indicate utterance types. Rows indicate card types. Critical trials are
marked in bold.}
\end{figure}

On critical trials, participants could derive implicatures in two ways.
First, on trials on which two animals were present on the card (e.g.,
cat and dog) but Bob guessed only one of them (e.g. \enquote{There is a
cat on the card}), the utterance could have a literal interpretation
(\enquote{There is a cat and possibly another animal on the card}) or an
exhaustive interpretation (\enquote{There is only a cat on the card}).
We refer to these trials as \enquote{exhaustive}. Second, on trials on
which two animals were on the card (e.g., a cat and a dog) and Bob used
a disjunciton (e.g., \enquote{There is a cat or a dog on the card}), the
utterance could have the literal, inclusive, interpretation, or a
pragmatic, exclusive interpretation. We refer to these trials as
\enquote{scalar}.

In order to assess the effect of the number of response options on
implicature rate, we manipulated number of response options in the
forced choice task between participants. We refer to the choice
conditions as \enquote{binary} (options: \emph{wrong}, \emph{right}),
\enquote{ternary} (options: \emph{wrong}, \emph{neither}, \emph{right}),
\enquote{quaternary} (options: \emph{wrong}, \emph{kinda wrong},
\emph{kinda right}, \emph{right}), and \enquote{quinary} (\emph{wrong},
\emph{kinda wrong}, \emph{neither}, \emph{kinda right}, \emph{right}).
Thus, the endpoint labels always remained the same. If there was an
uneven number of response options, the central option was
\emph{neither}. Participants were randomly assigned to one of the four
task conditions.

\subsection{Results and discussion}\label{results-and-discussion}

The collected dataset contains 50 participants in the binary task, 53 in
the ternary task, 43 in the quaternary task, and 54 in the quinary task.
Figures \ref{fig:binaryPlot} to \ref{fig:quinaryPlot} show the
proportions of response choices in each of the 8 trial types on each of
the four response tasks, respectively. We report the relevant patterns
of results qualitatively before turning to the quantitative analysis of
interest.

\subsubsection{Qualitative analysis}\label{qualitative-analysis}

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/binaryPlot-1.pdf}
\caption{\label{fig:binaryPlot}Proportion of responses for the binary forced
choice judgments. Error bars indicate 95\% bootstrapped confidence
intervals.}
\end{figure}

In the binary task, participants were at or close to ceiling in
responding \enquote{right} and \enquote{wrong} on unambiguously true and
false trials, respectively (see Figure \ref{fig:binaryPlot}). However,
on underinformative trials (i.e.~a \enquote{cat} or \enquote{cat or dog}
description for a card with both a cat and a dog), we observe pragmatic
behavior: on exhaustive trials, participants judged the utterance
\enquote{wrong} 14\% of the time; on scalar trials, participants judged
the utterance \enquote{wrong} 38\% of the time. That is, both under the
Weak and Strong link assumptions introduced in the Introduction,
inferred implicature rate on exhaustive trials is 14\% and on scalar
trials 38\%.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/ternaryPlot-1.pdf}
\caption{\label{fig:ternaryPlot}Proportion of responses for the ternary
forced choice judgments. Error bars indicate 95\% bootstrapped
confidence intervals.}
\end{figure}

In the ternary task, participants were also at or close to ceiling in
responding \enquote{right} and \enquote{wrong} on unambiguously true and
false trials, respectively (see Figure \ref{fig:ternaryPlot}). And
again, on underinformative trials (a \enquote{cat} and \enquote{cat or
dog} description for a card with both a cat and a dog), we observed
pragmatic behavior: on exhaustive trials, participants considered the
guess \enquote{wrong} 8\% of the time and neither wrong nor right 12\%
of the time. On scalar trials, participants judged the guess
\enquote{wrong} 23\% of the time and \enquote{neither} 11\% of the time.
This means that the Weak and Strong link lead to different conclusions
about implicature rates on the ternary task. Under the Weak link,
inferred implicature rate on exhaustive trials is 20\%; under the Strong
link it is only 8\%. Similarly, under the Weak link, inferred
implicature rate on scalar trials is 34\%; under the Strong link it is
only 23\%.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/quaternaryPlot-1.pdf}
\caption{\label{fig:quaternaryPlot}Proportion of responses for the
quaternary forced choice judgments. Error bars indicate 95\%
bootstrapped confidence intervals.}
\end{figure}

In the quaternary task (Figure \ref{fig:quaternaryPlot}), participants
were again at or close to ceiling in responding \enquote{right} and
\enquote{wrong} on 4 of the 6 unambiguously true and false trials.
However, with four response options, two of the control conditions
appear to be showing signs of pragmatic infelicity: when a conjunction
was used and only one of the animals was on the card, participants
considered the guess \enquote{wrong} most of the time (46\%), but they often
considered it \enquote{kinda wrong} (32\%) or even \enquote{kinda right} (19\%). This
suggests that perhaps participants considered the notion of a partially
true or correct statement in our experimental setting. Disjunctive
descriptions of cards with only one animal, while previously at ceiling
for \enquote{right} responses, were downgraded to only \enquote{kinda
right} 26\% of the time, presumably because these utterances are also
underinformative, though the degree of underinformativeness may be less
egregious than on scalar trials.

On underinformative exhaustive trials, we observed pragmatic behavior as
before: participants judged the guess \enquote{wrong} 2\% of the time,
\enquote{kinda wrong} 5\% of the time, and \enquote{kinda right} 66\% of
the time. On scalar trials, participants judged the guess
\enquote{wrong} 6\% of the time, \enquote{kinda wrong} 12\% of the time,
and \enquote{kinda right} 43\% of the times.

Thus, we are again forced to draw different conclusions about
implicature rates depending on whether we assume the Weak link or the
Strong link. Under the Weak link, inferred implicature rate on
exhaustive trials is 73\%; under the Strong link it is only 2\%.
Similarly, under the Weak link, inferred implicature rate on scalar
trials is 61\%; under the Strong link it is only 6\%.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/quinaryPlot-1.pdf}
\caption{\label{fig:quinaryPlot}Proportion of responses for the quinary
forced choice judgments. Error bars indicate 95\% bootstrapped
confidence intervals.}
\end{figure}

Finally, Figure \ref{fig:quinaryPlot} shows the proportion of responses
in the quinary task. Performance on the 4 pragmatically felicitous
control trials was again at floor and ceiling, respectively. The 2
control conditions in which the quaternary task had revealed pragmatic
infelicity again displayed that pragmatic infelicity in the quinary
task, suggesting that this is a robust type of pragmatic infelicity
that, nonetheless, requires fine-grained enough response options to be
detected experimentally.

On underinformative exhaustive trials, we observed pragmatic behavior as
before: participants judged the guess \enquote{wrong} 2\% of the time,
\enquote{kinda wrong} 1 and 1\% of the time, \enquote{neither} 1 and 1\%
of the time, and \enquote{kinda right} 72\% of the time. On scalar
trials, participants judged the guess \enquote{wrong} 6\% of the time,
\enquote{kinda wrong} 4\% of the time, \enquote{neither} 1\% of the
time, and \enquote{kinda right} 52\% of the time.

Thus, we would again draw different conclusions about implicature rates
depending on whether we assume the Weak link or the Strong link. Under
the Weak link, inferred implicature rate on exhaustive trials is 76 and
76\%; under the Strong link it is only 2\%. Similarly, under the Weak
link, inferred implicature rate on scalar trials is 63\%; under the
Strong link it is only 6\%.

\subsubsection{Quantitative analysis}\label{quantitative-analysis}

Our primary goal in this study was to test whether the estimated
implicature rate in the experimental task is affected by the linking
hypothesis and the number of response options available to participants.
To this end, we only analyzed the critical trials (exhaustive and
scalar). In particular, we classified each data point from critical
trials as constituting an implicature (1) or not (0) under the Strong
and Weak link. Figure \ref{fig:implicatureRatePlot} shows the resulting
implicature rates by condition and link. It is immediately apparent that there is variability in inferred implicature rate. In particular, the Weak link appears to result in greater estimates of
implicature rates in tasks with four or five response options, compared to the Strong link. For the binary and ternary task, the assumed link appears to play a much smaller role.

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/implicatureRatePlot-1.pdf}
\caption{\label{fig:implicatureRatePlot}Inferred implicature rates on
exhaustive and scalar trials as obtained with the binary, ternary,
quaternary, and quinary response task. Columns indicate link from
response to implicature rate (strong: proportion of \enquote{wrong
judgments; weak: proportion of non-'right} judgments).}
\end{figure}

\begin{table}

\caption{\label{tab:implicatureRate}\label{tab:modeltable}Model parameter estimates and their credible intervals. Rows marked with an asterisk in the evidence column do not contain 0 in the credible interval, thereby providing evidence for an effect.}
\centering
\begin{tabular}[t]{lrrrl}
\toprule
Predictors & Estimate & 2.5\% & 97.5\% & Evidence\\
\midrule
Intercept & -8.60 & -13.98 & -4.53 & *\\
Link = Weak & -0.15 & -4.86 & 4.77 & \\
Task = Quaternary & -1.83 & -8.08 & 4.20 & \\
Task = Quinary & -4.05 & -10.90 & 2.38 & \\
Task = Ternary & -1.45 & -7.31 & 4.56 & \\
\addlinespace
Implicature = Scalar & 6.09 & 1.00 & 12.29 & *\\
Link = Weak : Task = Quaternary & 14.03 & 7.24 & 21.88 & *\\
Link = Weak : Task = Quinary & 17.28 & 10.64 & 25.80 & *\\
Link = Weak : Task = Ternary & 3.81 & -1.49 & 9.22 & \\
Link = Weak : Implicature = Scalar & 0.90 & -4.01 & 6.43 & \\
\addlinespace
Task = Quaternary : Implicature = Scalar & -5.67 & -13.66 & 1.54 & \\
Task = Quinary : Implicature = Scalar & -2.31 & -9.30 & 4.61 & \\
Task = Ternary : Implicature = Scalar & -1.31 & -7.70 & 4.65 & \\
Link=Weak : Task=Quaternary : Implicature=Scalar & -3.29 & -12.07 & 4.55 & \\
Link=Weak : Task=Quinary : Implicature=Scalar & -7.74 & -16.59 & -0.16 & *\\
Link=Weak : Task=Ternary : Implicature=Scalar & -1.44 & -7.00 & 4.22 & \\
\bottomrule
\end{tabular}
\end{table}

To analyze the effect of link and response options on inferred
implicature rate, we used a Bayesian binomial mixed effects model using
the R packge \enquote{brms} (Bürkner \& others, 2016) with weakly informative priors.\footnote{For more information about the
  default priors of the \enquote{brms} package, see
  \href{ftp://cran.r-project.org/pub/R/web/packages/brms/brms.pdf}{the
  brms package manual}.} The model predicted the log odds of implicature
over no implicature from fixed effects of \emph{response type} (binary,
ternary, quaternary, quinary -- dummy-coded with binary as reference
level), \emph{link} (strong vs.~weak -- dummy-coded with strong as
reference level), and trial type (exhaustive vs.~scalar -- dummy-coded,
with exhaustive as reference level), as well as their two-way and
three-way interactions. Following Barr, Levy, Scheepers, and Tily
(2013), we included the maximal random effects structure justified by
the design: random intercepts for items (cards) and participants, random
by-participant slopes for link, trial type, and their interaction, and
random by-item slopes for link, trial type, response type, and their
interactions. Since the number of response options was a between-participant variable we did not include random slopes of response
options for participants. Four chains converged after 2000 iterations
each (warmup = 1000). Table \ref{tab:modeltable} summarizes the mean
parameter estimates and their 95\% credible intervals. \(\hat{R}=1\) for
all estimated parameters. All the analytical decisions described here
were pre-registered\footnote{Our preregistration can be accessed at
  \url{https://aspredicted.org/tq3sz.pdf}}.

The model provided evidence for the following effects: First, there was
a main effect of trial type such that scalar trials resulted in greater
implicature rates than exhaustive trials (Mean Estimate = 6.09, 95\%
Credible Interval={[}1, 12.29{]}). Second, there was an interaction
between link and number of response options such that the quaternary
task (Mean Estimate = 14.03, 95\% Credible Interval={[}7.24, 21.88{]})
and the quinary task (Mean Estimate = 17.28, 95\% Credible
Interval={[}10.64, 25.80{]}) resulted in greater
implicature rates with a weak link than with a strong link, but there was no evidence of a link-dependent difference in inferred implicature rate for the binary and ternary task. Finally, there was a three-way interaction between
link, trial type, and number of response options, driven by the binary/quinary contrast (Mean Estimate = -7.74,
95\% Credible Interval={[}-16.59, -0.16{]}). Simple effects analysis on only the binary and and quinary trials, separately for the exhaustive and scalar subset of the data, revealed that the three-way interaction is driven by a different effect of number of response options under the Weak vs Strong link for the two inference types. Specifically, on exhaustive trials, number of response options (2 vs.~5) only resulted in greater implicature rates under the Weak ($\beta$ = .2, $p <$ .0001), but not the Strong link ($\beta$ = -.8, $p <$ .82). In contrast, on scalar trials, number of response options (2 vs.~5) resulted in greater implicature rates under the Weak ($\beta$ = 3.6, $p <$ .005) link, and in lower implicature rates under the Strong link ($\beta$ = -4.0, $p <$ .0007).

In sum, both number of response options and link
affected the inferred implicature rate, as did the type of inference (exhaustive vs.~scalar).

\section{General Discussion}\label{general-discussion}

\subsection{Summary and methodological
discussion}\label{summary-and-methodological-discussion}

In this paper we asked whether linking hypothesis and number of
response options available to participants in truth value judgment tasks
affects inferred implicature rates. The results presented here suggest
they do. A linking assumption that considered the highest point on the
scale literal and any lower point pragmatic (Weak link) resulted in
higher implicature rates in tasks with 4 or 5 response options compared
to the standard two options. A linking hypothesis that considered the
lowest point on the scale pragmatic and any higher point literal (Strong
link) reported lower implicature rates in tasks with 4 or 5 options
compared to the standard two options. The results suggest that the
choice of linking hypothesis is a crucial analytical step that can
significantly impact the conclusions drawn from truth value judgment
tasks. In particular, there is danger for pragmatic ability to be both
under- and overestimated.

While the binary truth value judgement task avoids the analytic decision
between Strong and Weak linking hypothesis, the results reported here
suggest that binary tasks can also underestimate participants' pragmatic
competence. In binary tasks, participants are often given the lowest and
highest points on a scale (\enquote{wrong} vs. \enquote{right}) and are
asked to report pragmatic infelicities using the lowest point (e.g.
\enquote{wrong}). The study reported here showed that on trials with
true but pragmatically infelicitous descriptions, participants often
avoided the lowest point on the scale if they were given more
intermediate options. Even though the option \enquote{wrong} was
available to participants in all tasks, participants in tasks with
intermediate options chose it less often. In computing implicature rate,
this pattern manifested itself as a decrease in implicature rate under
the Strong link when more response options were provided, and an
increase in implicature rate under the Weak link when more response
options were provided. These observations are in line with Katsos and
Bishop (2011)'s argument that pragmatic violations are not as
severe as semantic violations and participants do not penalize them as
much. Providing participants with only the extreme ends of the scale
(e.g.~wrong/right, false/true) when pragmatic violations are considered
to be of an intermediate nature risks misrepresentation of participants'
pragmatic competence. It further suggests that in studies that use
binary tasks to investigate response-contingent processing, proportions
of \enquote{literal} responses may be a composite of both literal and
pragmatic underlying interpretations that just happen to get mapped
differently onto different response options by participants.

This study did not investigate the effect of response labels on the
inferred implicature rate. However, the results provided suggestive
evidence that some options better capture participant intuitions of
pragmatic infelicities than others. Among the intermediate options,
\enquote{kinda right} was chosen most often to report pragmatic
infelicities. The option \enquote{neither} was rarely used in the
ternary and quinary tasks (where it was used as a midpoint), suggesting
that participants interpreted pragmatic infelicities as different
degrees of being \enquote{right} and not \enquote{neither right nor
wrong.} Therefore, options that capture degrees of being \enquote{right}
like \enquote{kinda right} may prove most suitable for capturing
infelicity in the long run. We leave this as a methodological issue for
future research.

The study had three further design features worth investigating in
future work. First, the utterances were ostensibly produced by a
blindfolded character. This was an intentional decision to control for
violation of ignorance expectations with disjunction. A disjunction such
as \enquote{A or B} often carries an implication or expectation that the
speaker is not certain which alternative actually holds. Future work
should investigate how the violation of the ignorance expectation
interacts with link and number of response options in inferred
implicature rate. Second, in this study we considered exhaustive and
scalar implicatures with \emph{or}. If the observed effects of link and
number of response options hold in general, they should be observable
using other scales, e.g., on implicatures with \emph{some}. Finally, our
experiment was designed as a guessing game and the exact goal or
task-relevant Question Under Discussion of the game was left implicit.
Given the past literature on QUD effects on scalar implicature, we
expect that different goals -- e.g., to help the character win more
points vs.~to help the character be more accurate -- would affect how
strict or lenient participants are with their judgments and ultimately
affect implicature rate in the task (Degen \& Goodman, 2014; Zondervan,
2010). Future work should systematically vary the goal of the game and
explore its effects on the inferred implicature rate. But crucially,
it's unlikely that the observed effects of number of response options
and linking hypothesis on inferred implicature rate are dependent on any
of the discussed design choices.

\subsection{Revisiting the linking
hypothesis}\label{revisiting-linking-hypothesis}

On the traditional view of the link between implicature and behavior in
sentence verification tasks, scalar implicature is conceptualized as a
binary, categorical affair -- that is, an implicature is either
\enquote{calculated} or it isn't, and the behavioral reflexes of this
categorical interpretation process should be straightforwardly observed
in experimental paradigms. This assumption raises concerns for analyzing
variation in behavior on a truth value judgment task; for example, why
did the majority of respondents in the binary condition of our
experiment answer \enquote{right} to an utterance of the
underinformative \enquote{There is a cat or dog} when the card had both
a cat and a dog on it? And why did a sizeable minority nonetheless
choose \enquote{wrong} in this same condition?

To explain these data on the traditional view, we are forced to say that
a) not all participants calculated the implicature; or that b) some
participants who calculated the implicature did not choose the
anticipated (i.e., \enquote{wrong}) response due to some other cognitive
process which overrode the \enquote{correct} implicature behavior; or
some mixture of (a) and (b). We might similarly posit that one or both
of these factors underlie the variation in the ternary, quaternary, and
quinary conditions. However, without an understanding of how to
quantitatively specify the link between implicature calculation and its
behavioral expression, the best we can hope for on this approach is an
analysis which predicts general qualitative patterns in the data (e.g.~a
prediction of relatively more \enquote{right} responses than
\enquote{wrong} responses in a given trial of our binary truth value
judgment task, or a prediction of a rise in the rate of response of
\enquote{right}/\enquote{wrong} between two experimental conditions,
given some contextual manipulation). However, we should stress that to
the best of our knowledge, even a qualitative analysis of this kind of
variation in behavior on sentence verification tasks -- much less the
effect of the number of response choices on that behavior -- is largely
underdeveloped in the scalar implicature literature.

We contrast the above view of implicature and its behavioral reflexes
with an alternative linking hypothesis. Recent developments in the field
of probabilistic pragmatics have demonstrated that pragmatic production
and comprehension can be captured within the Rational Speech Act (RSA)
framework (Bergen, Levy, \& Goodman, 2016; Degen, Franke, \& Jäger,
2013; Degen, Tessler, \& Goodman, 2015; Frank \& Goodman, 2012; Franke
\& Jäger, 2016; Goodman \& Frank, 2016; Goodman \& Stuhlmüller, 2013;
Kao, Wu, Bergen, \& Goodman, 2014; Qing \& Franke, 2015). Much in the spirit of Gricean approaches to
pragmatic competence, the RSA framework takes as its point of departure
the idea that individuals are rational, goal-oriented communicative
agents, who in turn assume that their interlocutors similarly behave
according to general principles of cooperativity in communication. Just
as in more traditional Gricean pragmatics, pragmatic inference and
pragmatically-cooperative language production in the RSA framework are,
at their core, the product of counterfactual reasoning about alternative
utterances that one might produce (but does not, in the interest of
cooperativity). However, the RSA framework explicitly and quantitatively
models cooperative interlocutors as agents whose language production and
comprehension is a function of Bayesian probabilistic inference
regarding other interlocutors' expected behavior in a discourse context.

Specifically, in the RSA framework we model pragmatically competent
listeners as continuous probabilistic distributions over possible
meanings (states of the world) given an utterance which that listener
observes. The probability with which this listener \(L_1\) ascribes a
meaning \(s\) to an utterance \(u\) depends upon a prior probability
distribution of potential states of the world \(P_w\), and upon
reasoning about the communicative behavior of a speaker \(S_1\). \(S_1\)
in turn is modeled as a continuous probabilistic distribution over
possible utterances given an intended state of the world the speaker
intends to communicate. This distribution is sensitive to a rationality
parameter \(\alpha\), the production cost \(C\) of potential utterances,
and the informativeness of the utterance, quantified via a
representation of a literal listener \(L_0\) whose interpretation of an
utterance is in turn a function of that utterance's truth conditional
content \([[u]](s)\) and her prior beliefs about the state of the world
\(P_w(s)\).

\(P_{L_1}(s | u) \propto P_{S_1}(u | s) * P_w(s)\)

\(P_{S_1}(u | s) \propto exp(\alpha(log(L_0(s | u)) - C(u)))\)

\(P_{L_0}(s | u) \propto [[u]](s) * P_w(s)\)

This view contrasts with the traditional view in that it is rooted in a
quantitative formalization of pragmatic competence which provides us a
continuous measure of pragmatic reasoning. In the RSA framework,
individuals never categorically draw (or fail to draw) pragmatic
inferences about the utterances they hear. For example, exclusivity
readings of disjunction are represented in RSA as relatively lower
posterior conditional probability of a conjunctive meaning on the
\(P_L\) distribution given an utterance of \enquote{or}, compared to the
prior probability of that meaning. Thus, absent auxiliary assumptions
about what exactly would constitute \enquote{implicature}, it is not
even possible to talk about rate of implicature calculation in the RSA
framework. The upshot, as we show below, is that this view of pragmatic
competence does allow us to talk explicitly and quantitatively about
rates of observed behavior in sentence verification tasks.

We take inspiration from the RSA approach and treat participants'
behavior in our experimental tasks as the result of a soft-optimal
pragmatic speaker in the RSA framework. That is, following Degen and
Goodman (2014), we proceed on the assumption that behavior on sentence
verification tasks such as truth value judgment tasks, is best modeled
as a function of an individual's mental representation of a cooperative
speaker (\(S_1\) in the language of RSA) rather than of a pragmatic
listener who interprets utterances (\(P_{L_1}\)).\footnote{Degen and Goodman (2014) argue that sentence verification is more plausibly construed as a production task rather than as an interpretation task because participants, unlike in natural language comprehension, are provided with the ground truth about the state of the world that a speaker is describing. Thus, participants are in essence being asked to assess the quality of a speaker's utterance. In contrast, Degen and Goodman argue, true interpretation tasks are characterized by the listener inferring what the state of the world is that the speaker is describing, for instance by selecting from one of multiple interpretation options.} In their paper, Degen
\& Goodman show that sentence verification tasks are relatively more
sensitive to contextual features like the
Question Under Discussion than are sentence interpretation tasks, and
that this follows if sentence interpretation tasks -- but not sentence
verification tasks -- require an additional layer of counterfactual
reasoning about the intentions of a cooperative speaker.

A main desideratum of a behavioral linking hypothesis given the RSA view
of pragmatic competence is to transform continuous probability
distributions into categorical outputs (e.g.~responses of
\enquote{right}/''wrong'' in the case of the binary condition of our
experiment). For a given utterance \(u\) and an intended communicated
meaning \(s\), \(S_1\)(u \textbar{} s) outputs a conditional probability
of \(u\) given \(s\). For example, in the binary condition of our
experiment where a participant evaluated \enquote{There is a cat or a
dog} when there were both animals on the card, the participant has
access to the mental representation of \(S_1\) and hence to the \(S_1\)
conditional probability of producing the utterance \enquote{cat or dog}
given a dog and cat card: \(S_1\)(\enquote{cat or dog} \textbar{} cat
and dog). According to the linking hypothesis advanced here, the
participant provides a particular response to \(u\) if the RSA speaker
probability of \(u\) lies within a particular probability interval. We
model a responder, \(R\), who in the binary condition responds
\enquote{right} to an utterance \(u\) in world \(s\) just in case
\(S_1(u | s)\) exceeds some probability threshold \(\theta\):

R(u, w, \(\theta\))

= \enquote{right} iff \(S_1\)(u \textbar{} s) \(>\) \(\theta\)

= \enquote{wrong} otherwise

The model of a responder in the binary condition is extended intuitively
to the condition where participants had three response options. In this
case, we allow for two probability thresholds: \(\theta_1\), the minimum
standard for an utterance in a given world state to count as
\enquote{right}, and \(\theta_2\), the minimum standard for
\enquote{neither}. Thus, in the ternary condition, R(u, s, \(\theta_1\)
, \(\theta_2\)) is \enquote{right} iff \(S_1\)(u \textbar{} s)
\textgreater{} \(\theta_1\) and \enquote{neither} iff \(\theta_1\)
\textgreater{} \(S_1\)(u \textbar{} s) \textgreater{} \(\theta_2\). To
fully generalize the model to our five experimental conditions, we say
that \(R\) takes as its input an utterance \(u\), a world state \(s\),
and a number of threshold variables dependent on a variable \(c\),
corresponding to the experimental condition in which the participant
finds themself (e.g.~the range of possible responses available to
\(R\)).

Given c = \enquote{ternary}

R(u, w, \(\theta_1\) , \(\theta_2\))

= \enquote{right} iff \(S_1\)(u \textbar{} s) \(>\) \(\theta_1\)

= \enquote{neither} iff \(\theta_1\) \(>\) \(S_1\)(u \textbar{} s) \(>\)
\(\theta_2\)

= \enquote{wrong} otherwise

Given c = \enquote{quaternary}

R(u, w, \(\theta_1\) , \(\theta_2\), \(\theta_3\))

= \enquote{right} iff \(S_1\)(u \textbar{} s) \(>\) \(\theta_1\)

= \enquote{kinda right} iff \(\theta_1\) \(>\) \(S_1\)(u \textbar{} s)
\(>\) \(\theta_2\)

= \enquote{kinda wrong} iff \(\theta_2\) \(>\) \(S_1\)(u \textbar{} s)
\(>\) \(\theta_3\)

= \enquote{wrong} otherwise

Given c = \enquote{quinary}

R(u, w, \(\theta_1\) , \(\theta_2\), \(\theta_3\). \(\theta_4\))

= \enquote{right} iff \(S_1\)(u \textbar{} s) \(>\) \(\theta_1\)

=\enquote{kinda right} iff \(\theta_1\) \(>\) \(S_1\)(u \textbar{} s)
\(>\) \(\theta_2\)

= \enquote{neither} iff \(\theta_2\) \(>\) \(S_1\)(u \textbar{} s) \(>\)
\(\theta_3\)

= \enquote{kinda wrong} iff \(\theta_3\) \(>\) \(S_1\)(u \textbar{} s)
\(>\) \(\theta_4\)

= \enquote{wrong} otherwise

In an RSA model, \(S_1\)(u \textbar{} s) will be defined for any
possible combination of possible utterance and possible world state. One
consequence of this is that for the purposes of our linking hypothesis,
participants are modeled as employing the same decision criterion --
does \(S_1\)(u \textbar{} s) exceed the threshold? -- in both
\enquote{implicature} and \enquote{non-implicature} conditions of a
truth value judgment task experiment. That is, participants never
evaluate utterances directly on the basis of logical truth or falsity:
for example, our blindfolded character Bob's guess of \enquote{cat and
dog} on a cat and dog card trial is \enquote{right} to the vast majority
of participants not because the guess is logically true but because
\(S_1\)(\enquote{cat and dog} \textbar{} cat and dog) is exceedingly
high.

For further illustration, we use our definition of a
pragmatically-competent speaker \(S_1\) (as defined above) to calculate
the speaker probabilities of utterances in states of the world
corresponding to our experimental conditions (i.e., for \enquote{cat},
\enquote{dog}, \enquote{cat and dog}, and \enquote{elephant}, given
either a cat on the card, or both a cat and a dog on the card). In
calculating these probabilities, we assume that the space of possible
utterances is the set of utterances made by Bob in our experiment
(i.e.~any possible single, disjunctive, or conjunctive guess involving
\enquote{cat}, \enquote{dog}, or \enquote{elephant}). For the purposes
of our model, we assume a uniform cost term on all utterances. We
furthermore assume that the space of possible meanings corresponds to
the set of possible card configurations that a participant may have seen
in our experiment, and that the prior probability distribution over
these world states is uniform. Lastly, we set \(\alpha\) -- the speaker
rationality parameter -- to 1. The resulting speaker probabilities are
shown in Figure \ref{fig:speakerprobs}.\footnote{Note that the
  probabilities in each facet don't sum to 1 because the model considers
  all possible disjunctive, conjunctive, and simple utterances, while we
  are only visualizing the ones corresponding to the experimental
  conditions.}

\begin{figure}
\centering
\includegraphics{writeup_files/figure-latex/speakerprobs-1.pdf}
\caption{\label{fig:speakerprobs}Speaker probabilities of utterances on the
exhaustive and scalar trials, as obtained using the model described in
this section.}
\end{figure}

The linking hypothesis under discussion assumes that speaker
probabilities of utterance given meaning are invariant across a) our
four different experimental conditions, b) across participants, and c)
within participants (that is, participants do not update
their \(S_1\) distribution in a local discourse context). We note that
the assumption (b) may conceivably be relaxed by allowing one or more of
the parameters in the model -- including the prior probability over
world states \(P_w\), the cost function on utterances \(C\), or the
rationality parameter \(\alpha\) -- to vary across participants. We also
note that assumption (c) in particular is in tension with a growing body
of empirical evidence that semantic and pragmatic interpretation is
modulated by rapid adaptation to the linguistic and social features of
one's interlocutors (Fine, Jaeger, Farmer, \& Qian, 2013; Kleinschmidt
\& Jaeger, 2015; Yildirim, Degen, Tanenhaus, \& Jaeger, 2016).

However, if we should like to keep the above simplifying assumptions in place, then
this linking hypothesis commits us to explaining variation in the data in terms
of the threshold parameters of our responder model \(R\). Consider first
the variation in response across different experimental conditions on a
given trial, e.g.~evaluation of a guess of \enquote{cat and dog} when
the card contains both a cat and a dog. The variation in the proportion
of responses of \enquote{right} on this trial between the binary,
ternary, quaternary, and quinary conditions indicates that the threshold
value for \enquote{right} responses must vary across conditions; that
is, we predict that the \(\theta\) of the binary condition will differ
from, e.g., the \(\theta_1\) of the ternary condition as well as the
\(\theta_1\) of the quaternary condition. We also observed variation in
response on this trial within a single condition (for example, a
sizeable minority of participants responded \enquote{wrong} to this
trial in the binary condition). Thus, this linking hypothesis is
committed to the notion that threshold values may vary across
participants, such that a speaker probability of utterance \(S_1\)(u
\textbar{} s) can fall below \(\theta\) for some subset of participants
while \(S_1\)(u \textbar{} s) itself remains constant across
participants.

Lastly, for two utterances of the same
conditional probability and in the same experimental condition, 
participants in our experiment sometimes provided a judgment of
\enquote{right} to one utterance but \enquote{wrong} to the other. That
is, there was within-subject variation in this
experiment. One way to represent such variation would be to posit that
the parameterization of threshold values proceeds stochastically and
that threshold values are recalibrated for every individual sentence
verification task. Rather than representing a threshold as a discrete
value N between 0 and 1, we can represent that threshold as a
distribution over possible threshold values -- with mass centered around
N. Whenever an individual encounters  a single trial of our truth value judgment task experiment, a
threshold value is sampled from this distribution. By allowing values of \(\theta\) to vary stochastically in this way, we can capture that \(S_1\)(u \textbar{} s) can fall both above and below \(\theta\) for a given
participant.

The model in its present form already captures an interesting asymmetry in inferred implicature rates between exhaustive and scalar trials of the experiment: note specifically (c.f. Figure 8) that inferred implicature rates are greater in the binary and ternary conditions for scalar trials over exhaustive trials. This is expected given the model's inferred speaker probabilities: the speaker probability of producing ``There is a cat on the card'' in the context of there being a cat and dog on the card (an exhaustive implicature-inducing trial) is greater than the speaker probability of producing ``There is a cat or a dog on the card'' in that same context (a scalar implicature-inducing trial). Assuming noisy  $\theta$ values centered around N, participants are expected to respond `Right' more frequently on exhaustive than on scalar trials, which is precisely what is observed. Recall that these probabilities were derived via the simplifying assumption of uniform cost on utterances; in fact, adding cost to relatively complex disjunctive sentences over simple declarative sentences only predicts a more pronounced asymmetry in the experimentally-observed direction.

%\change{While we have offered only a preliminary qualitative sketch of the model's behavior here, future work should systematically analyze the quantitative fit between model predictions and the data. }

%\change{Less readily explained is why this asymmetry disappears with the introduction of more response options, i.e. in the quaternary and quinary conditions. If anything, the opposite numerical trend is observed in these conditions: that is, inferred implicature rates are slightly greater in exhaustive trials over scalar trials. Future work should investigate the interaction between trial type (exhaustive vs. scalar) and number of response options.}

\change{As suggested above, the quantitative predictions of our model will depend crucially on the values assigned to its free parameters - including (but not limited to) the probability thresholds and speaker costs of possible utterances. However, the values of these parameters can be estimated in a principled and informed manner through Bayesian statistical analysis of our experimental data. Samples from prior distributions over possible parameter values yield predicted patterns of response, which are then compared against empirically-observed response patterns in order to determine the a posteriori probability that these values are in fact the 'real' latent parameter values. The resulting posterior distributions are sampled from in turn, in order to parameterize the model and assess overall quantitative fit given the data. Though we leave a quantitative assessment of our model to future work (Waldon et al., in prep), we sketch the general procedure here to emphasize that the model is amenable to rigorous and data-driven evaluation.}

One empirical problem is the pattern of responses we observed
for \enquote{cat and dog} on trials where there was only a cat on the
card. Because this utterance is strictly false in this world state, it
is surprising -- on both the traditional view as well as on the account
developed here -- that participants assigned this utterance ratings
above \enquote{wrong} with any systematicity. However, this is 
what we observed, particulary in the quaternary and quinary conditions
of the experiment, where a sizeable minority of participants considered
this utterance \enquote{kinda right}. As Figure \ref{fig:speakerprobs}
demonstrates, the conditional speaker probability of this utterance in
this world state is 0; thus, there is no conceivable threshold value
that would allow this utterance to ever be rated above \enquote{wrong}
(on the reasonable assumption that the thresholds in our responder model
\(R\) should be nonzero). Any linking hypothesis will have to engage
with this data point, and we leave to future work an analysis which
captures participants' behavior in this condition.

For the time being, however, we present the above analysis as a proof of
concept for the following idea: by relaxing the assumptions of the
traditional view of scalar implicature---namely, that scalar implicatures
either are or are not calculated, and that behavior on sentence
verification tasks directly reflects this binary interpretation
process---we can propose quantitative models of the variation in
behavior that is observed in experimental settings. We note that the linking
hypothesis proposed here is just one in the space of possible hypotheses. For
example, one might reject this threshold-based analysis in favor of one
whereby responses are the outcomes of sampling on the (pragmatic speaker
or pragmatic listener) probability distributions provided by an RSA
model. We leave this systematic, quantitative investigation to future work. For now we
emphasize that explicit computational modeling of behavioral responses is a tool that is  available to researchers in experimental
pragmatics. While using the RSA framework as the modeling tool requires revising traditional assumptions about the nature of scalar
implicature by relaxing the crisp notion of scalar
implicature as something that is or is not \enquote{calculated} in
interpretation, it provides new flexibility to explicitly discuss
 behavior in experimental settings. One need not adopt the RSA framework as the tool for hypothesizing and testing the link between theoretical constructs and behavior in pragmatic experiments. However, the empirical findings we have reported here---that the inferences researchers draw about ``implicature rate'' are volatile and depend on various features of the paradigm and the linking hypothesis employed--- strongly suggest that experimental pragmatics  as a field must engage more seriously with the foundational questions of what we are measuring in the experiments we run.

Concluding, we have shown in this paper that inferred
\enquote{implicature rate} -- a ubiquitous notion in theoretical and experimental
pragmatics -- as estimated in truth value judgment tasks, depends on
both the number of responses participants are provided with as well as
on the linking hypothesis from proportion of behavioral responses to
\enquote{implicature rate}. We further sketched an alternate linking
hypothesis that treats behavioral responses as the result of
probabilistic reasoning about speakers' likely productions. While a
thorough model comparison is still outstanding, this kind of linking
hypothesis opens a door towards more systematic and rigorous formulation
and testing of linking hypotheses between theoretical notions of
interest in pragmatics and behavioral responses in experimental
paradigms.

\section{Author Responsibilities}
All authors contributed to the conception and design of the study. MJ conducted the online survey studies; reported the results and performed the statistical analysis; BW conducted the modeling and wrote the discussion section; JD wrote the theoretical introduction, and contributed to the experimental section and the discussion section and the modeling sections. All authors contributed to manuscript revision, read and approved the submitted version.

\section{References}\label{references}

\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-Barner2011}{}
Barner, D., Brooks, N., \& Bale, A. (2011). Accessing the unsaid: the
role of scalar alternatives in children's pragmatic inference.
\emph{Cognition}, \emph{118}(1), 84--93.
doi:\href{https://doi.org/10.1016/j.cognition.2010.10.010}{10.1016/j.cognition.2010.10.010}

\hypertarget{ref-barr2013random}{}
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random
effects structure for confirmatory hypothesis testing: Keep it maximal.
\emph{Journal of Memory and Language}, \emph{68}(3), 255--278.

\hypertarget{ref-BenzGotzner2014}{}
Benz, A., \& Gotzner, N. (2014). Embedded implicatures revisited: Issues
with the Truth-Value Judgment Paradigm. In J. Degen, M. Franke, \& N. D.
Goodman (Eds.), \emph{Proceedings of the formal \& experimental
pragmatics workshop, European Summer School for Language, Logic and
Information (ESSLLI)} (pp. 1--6). Tübingen.

\hypertarget{ref-Bergen2012}{}
Bergen, L., \& Grodner, D. J. (2012). Speaker knowledge influences the
comprehension of pragmatic inferences. \emph{Journal of Experimental
Psychology. Learning, Memory, and Cognition}, \emph{38}(5), 1450--60.
doi:\href{https://doi.org/10.1037/a0027850}{10.1037/a0027850}

\hypertarget{ref-Bergen2016}{}
Bergen, L., Levy, R., \& Goodman, N. (2016). Pragmatic reasoning through
semantic inference. \emph{Semantics and Pragmatics}, \emph{9}(1984),
1--46. doi:\href{https://doi.org/10.3765/sp.9.20}{10.3765/sp.9.20}

\hypertarget{ref-Bonnefon2009}{}
Bonnefon, J.-F., Feeney, A., \& Villejoubert, G. (2009). When some is
actually all: scalar inferences in face-threatening contexts.
\emph{Cognition}, \emph{112}(2), 249--58.
doi:\href{https://doi.org/10.1016/j.cognition.2009.05.005}{10.1016/j.cognition.2009.05.005}

\hypertarget{ref-Bott2016}{}
Bott, L., \& Chemla, E. (2016). Shared and distinct mechanisms in
deriving linguistic enrichment. \emph{Journal of Memory and Language},
\emph{91}, 117--140.
doi:\href{https://doi.org/10.1016/j.jml.2016.04.004}{10.1016/j.jml.2016.04.004}

\hypertarget{ref-Bott2004}{}
Bott, L., \& Noveck, I. (2004). Some utterances are underinformative:
The onset and time course of scalar inferences. \emph{Journal of Memory
and Language}, \emph{51}(3), 437--457.
doi:\href{https://doi.org/10.1016/j.jml.2004.05.006}{10.1016/j.jml.2004.05.006}

\hypertarget{ref-Breheny2013}{}
Breheny, R., Ferguson, H. J., \& Katsos, N. (2013). Taking the epistemic
step: Toward a model of on-line access to conversational implicatures.
\emph{Cognition}, \emph{126}(3), 423--40.
doi:\href{https://doi.org/10.1016/j.cognition.2012.11.012}{10.1016/j.cognition.2012.11.012}

\hypertarget{ref-Breheny2006}{}
Breheny, R., Katsos, N., \& Williams, J. (2006). Are generalised scalar
implicatures generated by default? An on-line investigation into the
role of context in generating pragmatic inferences. \emph{Cognition},
\emph{100}(3), 434--63.
doi:\href{https://doi.org/10.1016/j.cognition.2005.07.003}{10.1016/j.cognition.2005.07.003}

\hypertarget{ref-burkner2016brms}{}
Bürkner, P.-C., \& others. (2016). brms: An R package for bayesian
multilevel models using Stan. \emph{Journal of Statistical Software},
\emph{80}(1), 1--28.

\hypertarget{ref-Chemla2011}{}
Chemla, E., \& Spector, B. (2011). Experimental evidence for embedded
scalar implicatures. \emph{Journal of Semantics}, \emph{28}(3),
359--400.

\hypertarget{ref-chierchia2001}{}
Chierchia, G., Crain, S., Teresa, M., Guasti, M. T., Gualmini, A., \&
Meroni, L. (2001). The acquisition of disjunction: Evidence for a
grammatical view of scalar implicatures. In Anna H.-J. Do Laura
Domínguez \& A. Johansen (Eds.), \emph{Proceedings of the 25th annual
boston university conference on language development} (pp. 157--168).
Somerville, MA: Cascadilla Press.

\hypertarget{ref-DeNeys2007}{}
De Neys, W., \& Schaeken, W. (2007). When people are more logical under
cognitive load - dual task impact on scalar implicature.
\emph{Experimental Psychology}, \emph{54}(2), 128--133.
doi:\href{https://doi.org/10.1027/1618-3169.54.2.128}{10.1027/1618-3169.54.2.128}

\hypertarget{ref-Degen2015}{}
Degen, J. (2015). Investigating the distribution of ``some'' (but not
``all'') implicatures using corpora and web-based methods.
\emph{Semantics and Pragmatics}, \emph{8}(11), 1--55.
doi:\href{https://doi.org/10.3765/sp.8.11}{10.3765/sp.8.11}

\hypertarget{ref-Degen2014}{}
Degen, J., \& Goodman, N. D. (2014). Lost your marbles? The puzzle of
dependent measures in experimental pragmatics. In P. Bello, M. Guarini,
M. McShane, \& B. Scassellati (Eds.), \emph{Proceedings of the 36th
annual conference of the cognitive science society} (pp. 397--402).

\hypertarget{ref-DegenTanenhaus2015}{}
Degen, J., \& Tanenhaus, M. K. (2015). Processing scalar implicature A
constraint-based approach. \emph{Cognitive Science}, \emph{39}(4),
667--710.
doi:\href{https://doi.org/10.1111/cogs.12171}{10.1111/cogs.12171}

\hypertarget{ref-DegenTanenhaus2016}{}
Degen, J., \& Tanenhaus, M. K. (2016). Availability of alternatives and
the processing of scalar implicatures: A visual world eye-tracking
study. \emph{Cognitive Science}, \emph{40}(1), 172--201.
doi:\href{https://doi.org/10.1111/cogs.12227}{10.1111/cogs.12227}

\hypertarget{ref-Degen2013}{}
Degen, J., Franke, M., \& Jäger, G. (2013). Cost-based pragmatic
inference about referential expressions. \emph{Proceedings of the 35th
Annual Conference of the Cognitive Science Society}, 376--281.

\hypertarget{ref-DegenTG2015}{}
Degen, J., Tessler, M. H., \& Goodman, N. D. (2015). Wonky worlds:
Listeners revise world knowledge when utterances are odd.
\emph{Proceedings of the 37th Annual Conference of the Cognitive Science
Society}, (2), 548--553.

\hypertarget{ref-Doran2012}{}
Doran, R., Ward, G., Larson, M., McNabb, Y., \& Baker, R. E. (2012). A
novel experimental paradigm for distinguishing between what is said and
what is implicated. \emph{Language}, \emph{88}, 124--154.

\hypertarget{ref-Fine2013}{}
Fine, A. B., Jaeger, T. F., Farmer, T. F., \& Qian, T. (2013). Rapid
expectation adaptation during syntactic comprehension. \emph{PLoS ONE},
\emph{8}(10).
doi:\href{https://doi.org/10.1371/\%20journal.pone.0077661}{10.1371/ journal.pone.0077661}

\hypertarget{ref-Frank2012}{}
Frank, M. C., \& Goodman, N. D. (2012). Predicting pragmatic reasoning
in language games. \emph{Science}, \emph{336}, 998.

\hypertarget{ref-Franke2016}{}
Franke, M., \& Jäger, G. (2016). Probabilistic pragmatics, or why Bayes'
rule is probably important for pragmatics. \emph{Zeitschrift Für
Sprachwissenschaft}, \emph{35}(1), 3--44.

\hypertarget{ref-Geurts2010}{}
Geurts, B. (2010). \emph{Quantity implicatures}. Cambridge: Cambridge
Univ Press.

\hypertarget{ref-Geurts2009}{}
Geurts, B., \& Pouscoulous, N. (2009). Embedded implicatures?!?
\emph{Semantics and Pragmatics}, \emph{2}, 1--34.
doi:\href{https://doi.org/10.3765/sp.2.4}{10.3765/sp.2.4}

\hypertarget{ref-Goodman2016}{}
Goodman, N. D., \& Frank, M. C. (2016). Pragmatic language
interpretation as probabilistic inference. \emph{Trends in Cognitive
Sciences}, \emph{20}(11), 818--829.
doi:\href{https://doi.org/10.1016/j.tics.2016.08.005}{10.1016/j.tics.2016.08.005}

\hypertarget{ref-Goodman2013}{}
Goodman, N. D., \& Stuhlmüller, A. (2013). Knowledge and implicature:
modeling language understanding as social cognition. \emph{Topics in
Cognitive Science}, \emph{5}(1), 173--84.
doi:\href{https://doi.org/10.1111/tops.12007}{10.1111/tops.12007}

\hypertarget{ref-grice1975}{}
Grice, H. P. (1975). Logic and conversation. \emph{Syntax and
Semantics}, \emph{3}, 41--58.

\hypertarget{ref-Grodner2010}{}
Grodner, D. J., Klein, N. M., Carbary, K. M., \& Tanenhaus, M. K.
(2010). ``Some,'' and possibly all, scalar inferences are not delayed:
Evidence for immediate pragmatic enrichment. \emph{Cognition},
\emph{116}(1), 42--55.
doi:\href{https://doi.org/10.1016/j.cognition.2010.03.014}{10.1016/j.cognition.2010.03.014}

\hypertarget{ref-Hirschberg1985}{}
Hirschberg, J. (1985). \emph{A Theory of Scalar Implicature} (PhD thesis
No. MS-CIS-85-56). University of Pennsylvania; Garland Publishing
Company.

\hypertarget{ref-Horn1972}{}
Horn, L. (1972). \emph{On the Semantic Properties of the Logical
Operators in English} (PhD thesis). UCLA.

\hypertarget{ref-horn1984}{}
Horn, L. (1984). Toward a new taxonomy for pragmatic inference: Q-based
and R-based implicature. In D. Schiffrin (Ed.), \emph{Meaning, form, and
use in context: Linguistic applications} (pp. 11--42). Washington:
Georgetown University Press.

\hypertarget{ref-Horowitz2017}{}
Horowitz, A. C., Schneider, R. M., \& Frank, M. C. (2017). The trouble
with quantifiers: Exploring children's deficits in scalar implicature.
\emph{Child Development}, 1--40.
doi:\href{https://doi.org/10.1111/cdev.13014}{10.1111/cdev.13014}

\hypertarget{ref-huang2009}{}
Huang, Y. T., \& Snedeker, J. (2009). On-line interpretationf of scalar
quantifiers: Insight into the semantics-pragmatics interface.
\emph{Cognitive Psychology}, \emph{58}, 376--415.

\hypertarget{ref-Kao2014}{}
Kao, J., Wu, J., Bergen, L., \& Goodman, N. D. (2014). Nonliteral
understanding of number words. \emph{Proceedings of the National Academy
of Sciences of the United States of America}, \emph{111}(33),
12002--12007.
doi:\href{https://doi.org/10.1073/pnas.1407479111}{10.1073/pnas.1407479111}

\hypertarget{ref-Katsos2011}{}
Katsos, N., \& Bishop, D. V. M. (2011). Pragmatic tolerance:
implications for the acquisition of informativeness and implicature.
\emph{Cognition}, \emph{120}(1), 67--81.
doi:\href{https://doi.org/10.1016/j.cognition.2011.02.015}{10.1016/j.cognition.2011.02.015}

\hypertarget{ref-Kleinschmidt2015}{}
Kleinschmidt, D. F., \& Jaeger, T. F. (2015). Robust speech perception:
Recognize the familiar, generalize to the similar, and adapt to the
novel. \emph{Psychological Review}, \emph{122}(2), 148--203.
doi:\href{https://doi.org/10.1037/a0038695}{10.1037/a0038695}

\hypertarget{ref-levinson2000}{}
Levinson, S. C. (2000). \emph{Presumptive Meanings - The Theory of
Generalized Conversational Implicature}. MIT Press.

\hypertarget{ref-DeMarneffe2017}{}
Marneffe, M.-C. de, \& Tonhauser, J. (2016). Inferring meaning from
indirect answers to polar questions: The contribution of the
rise-fall-rise contour. In E. Onea, M. Zimmermann, \& K. von Heusinger
(Eds.), \emph{Questions in discourse}. Leiden: Brill Publishing.

\hypertarget{ref-Musolino2004}{}
Musolino, J. (2004). The semantics and acquisition of number words:
integrating linguistic and developmental perspectives. \emph{Cognition},
\emph{93}(1), 1--41.
doi:\href{https://doi.org/10.1016/j.cognition.2003.10.002}{10.1016/j.cognition.2003.10.002}

\hypertarget{ref-Noveck2001}{}
Noveck, I. (2001). When children are more logical than adults:
experimental investigations of scalar implicature. \emph{Cognition},
\emph{78}(2), 165--188.

\hypertarget{ref-noveck2008}{}
Noveck, I. A., \& Reboul, A. (2008). Experimental pragmatics: a Gricean
turn in the study of language. \emph{Trends in Cognitive Sciences},
\emph{12}(11), 425--431.
doi:\href{https://doi.org/10.1016/j.tics.2008.07.009}{10.1016/j.tics.2008.07.009}

\hypertarget{ref-Noveck2003}{}
Noveck, I., \& Posada, A. (2003). Characterizing the time course of an
implicature: An evoked potentials study. \emph{Brain and Language},
\emph{85}(2), 203--210.
doi:\href{https://doi.org/10.1016/S0093-934X(03)00053-1}{10.1016/S0093-934X(03)00053-1}

\hypertarget{ref-Papafragou2004}{}
Papafragou, A., \& Tantalou, N. (2004). Children's computation of
implicatures. \emph{Language Acquisition}, \emph{12}(1), 71--82.

\hypertarget{ref-Papafragou2003}{}
Papafragou, A., \& Musolino, J. (2003). Scalar implicature: experiments at the semantics-pragmatics interface. \emph{Cognition}, \emph{86}(3), 253--282.

\hypertarget{ref-Politzer-Ahles2013}{}
Politzer-Ahles, S., \& Fiorentino, R. (2013). The Realization of Scalar
Inferences: Context Sensitivity without Processing Cost. \emph{PLoS
ONE}, \emph{8}(5).
doi:\href{https://doi.org/10.1371/journal.pone.0063943}{10.1371/journal.pone.0063943}

\hypertarget{ref-Potts2015}{}
Potts, C., Lassiter, D., Levy, R., \& Frank, M. C. (2015). Embedded
implicatures as pragmatic inferences under compositional lexical
uncertainty. \emph{Journal of Semantics}, \emph{33}(1975), 755--802.
doi:\href{https://doi.org/10.1093/jos/ffv012}{10.1093/jos/ffv012}

\hypertarget{ref-Qing2015}{}
Qing, C., \& Franke, M. (2015). Variations on a Bayesian theme:
Comparing Bayesian models of referential reasoning. In H. Zeevat \&
H.-C. Schmitz (Eds.), \emph{Bayesian natural language semantics and
pragmatics} (Vol. 2, pp. 201--220). Cham, Switzerland: Springer
International Publishing.
doi:\href{https://doi.org/10.1007/978-3-319-17064-0_9}{10.1007/978-3-319-17064-0\_9}

\hypertarget{ref-sauerland2004scalar}{}
Sauerland, U. (2004). Scalar implicatures in complex sentences.
\emph{Linguistics and Philosophy}, \emph{27}(3), 367--391.

\hypertarget{ref-Scontras2017}{}
Scontras, G., Degen, J., \& Goodman, N. D. (2017). Subjectivity predicts
adjective ordering preferences. \emph{Open Mind: Discoveries in
Cognitive Science}, \emph{1}(1), 53--65.
doi:\href{https://doi.org/10.1162/opmi}{10.1162/opmi}

\hypertarget{ref-Sikos2019}{}
Sikos, L., Kim, M., \& Grodner, D.J. (in press). Social context modulates tolerance for pragmatic violations in binary but not graded judgments. \emph{Frontiers in Psychology}.

\hypertarget{ref-Stiller2015}{}
Stiller, A. J., Goodman, N. D., \& Frank, M. C. (2015). Ad-hoc
implicature in preschool children. \emph{Language Learning and
Development}, \emph{11}(2), 176--190.
doi:\href{https://doi.org/10.1080/15475441.2014.927328}{10.1080/15475441.2014.927328}

\hypertarget{ref-Tanenhaus2004}{}
Tanenhaus, M. K. (2004). On-line sentence processing: past, present and
future. In M. Carreiras \& C. Clifton (Eds.), \emph{On-line sentence
processing: ERPS, eye movements and beyond} (pp. 371--392). London, UK:
Psychology Press.

\hypertarget{ref-Tomlinson2013}{}
Tomlinson, J. M., Bailey, T. M., \& Bott, L. (2013). Possibly all of
that and then some: Scalar implicatures are understood in two steps.
\emph{Journal of Memory and Language}, \emph{69}(1), 18--35.
doi:\href{https://doi.org/10.1016/j.jml.2013.02.003}{10.1016/j.jml.2013.02.003}

\hypertarget{ref-VanTiel2014}{}
van Tiel, B., van Miltenburg, E., Zevakhina, N., \& Geurts, B. (2016).
Scalar diversity. \emph{Journal of Semantics}.
doi:\href{https://doi.org/10.1093/jos/ffu017}{10.1093/jos/ffu017}

\hypertarget{ref-Yildirim2016}{}
Yildirim, I., Degen, J., Tanenhaus, M.K., \& Jaeger, T.F. (2016). Talker-specificity and adaptation in quantifier interpretation. \emph{Journal of Memory and Language}, \emph{87}, pp. 128-143.

\hypertarget{ref-Zondervan2010}{}
Zondervan, A. (2010). \emph{Scalar implicatures or focus: an
experimental approach} (PhD thesis). Universiteit Utrecht, Amsterdam.

\end{document}